### 2019-02-20 manuscript/supporting_info.tex
need to make this a multi-page figure

\documentclass[12pt]{article}  
\usepackage{amsmath}
\usepackage{url}
\usepackage[dvips]{graphicx}
\usepackage{multirow}
\usepackage{geometry}
\usepackage{pdflscape}
\usepackage{gensymb}
% \usepackage{rotating}
% make Figure 1 etc bold
\usepackage[labelfont=bf]{caption}
\usepackage{setspace}

\usepackage[running]{lineno}

\usepackage{dcolumn}
\newcolumntype{d}[1]{D{.}{.}{#1}}

%\usepackage{overcite}
\usepackage[round]{natbib}

\newcommand{\expect}[1]{\left\langle #1 \right\rangle}
\newcommand{\etal}{\textit{et al.\ }}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

% the abstract formatting
\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}
\renewcommand\refname{References}

% margin sizes`
\topmargin 0.0cm
\oddsidemargin 0.2cm

\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\title{Conservation of interaction partners between related plants varies widely across communities and between plant families - Supporting Information}

\author{Alyssa R. Cirtwill$^{1}$, Giulio V. Dalla Riva$^{2}$, Nick J. Baker$^{1}$,\\
 Joshua A. Thia$^{1,3}$, Christie J. Webber$^{1}$, Daniel B. Stouffer$^{1}$, MSC helpers from Linkoping}
\date{\small$^1$Centre for Integrative Ecology, School of Biological Sciences\\
    \medskip$^2$Biomathematics Research Centre, School of Mathematics and Statistics\\
            University of Canterbury\\Private Bag 4800\\
Christchurch 8140, New Zealand\\
\medskip$^3$Present Address: School of Biological Sciences\\
University of Queensland\\Brisbane, QLD 4072, Australia }

\begin{document}
\maketitle
\baselineskip=8.5mm
\begin{spacing}{1.0}

\clearpage

\section*{Supporting information 1: Sources for networks}

  \begin{table}[!h]
    \caption{Original sources for each network used in our analyses.}
    \label{sources}
    \begin{tabular}{|l l l |}
    \hline
    Network & Network type & Original source \\
    \hline
    M\_PL\_001  & Pollination & \citep{Arroyo1982}  \\
    M\_PL\_002  & Pollination & \citep{Arroyo1982}  \\
    M\_PL\_003  & Pollination & \citep{Arroyo1982}  \\
    M\_PL\_004  & Pollination & \citep{Barrett1987} \\
    M\_PL\_005  & Pollination & \citep{Clements1923}  \\
    M\_PL\_006  & Pollination & \citep{Dicks2002} \\
    M\_PL\_007  & Pollination & \citep{Dicks2002} \\
    M\_PL\_008  & Pollination & \citep{Dupont2003}  \\
    M\_PL\_009  & Pollination & \citep{Elberling1999} \\
    M\_PL\_010  & Pollination & Elberling, H. \& Olesen, J. M. Unpublished. \\
    M\_PL\_011  & Pollination & \citep{Olesen2002a}  \\
    M\_PL\_012  & Pollination & Olesen, J. M. Unpublished.  \\
    M\_PL\_013  & Pollination & \citep{Ollerton2003}  \\
    M\_PL\_014  & Pollination & \citep{Hocking1968} \\
    M\_PL\_015  & Pollination & \citep{Petanidou1991} \\
    M\_PL\_016  & Pollination & \citep{Herrera1988} \\
    M\_PL\_017  & Pollination & \citep{Memmott1999} \\
    M\_PL\_018  & Pollination & Olesen, J. M. Unpublished.  \\
    M\_PL\_019  & Pollination & \citep{Inouye1988}  \\
    M\_PL\_020  & Pollination & \citep{Kevan1970} \\
    M\_PL\_021  & Pollination & \citep{Kato1999}  \\
    M\_PL\_022  & Pollination & \citep{Medan2002} \\
    M\_PL\_023  & Pollination & \citep{Medan2002} \\
    M\_PL\_024  & Pollination & \citep{Mosquin1967} \\
    M\_PL\_025  & Pollination & \citep{Motten1982}  \\
    M\_PL\_026  & Pollination & \citep{McMullen1993}  \\
    M\_PL\_027  & Pollination & \citep{Primack1983} \\
    M\_PL\_028  & Pollination & \citep{Primack1983} \\
    M\_PL\_029  & Pollination & \citep{Primack1983} \\
    M\_PL\_030  & Pollination & \citep{Ramirez1992} \\
    M\_PL\_031  & Pollination & \citep{Ramirez1989} \\
    M\_PL\_032  & Pollination & \citep{Schemske1978}  \\
    M\_PL\_033  & Pollination & \citep{Small1976} \\
    M\_PL\_034  & Pollination & \citep{SmithRamirez2005}  \\
    M\_PL\_035  & Pollination & \citep{Percival1974}  \\
    M\_PL\_036  & Pollination & Olesen, J. M. Unpublished.  \\
    \hline
    \end{tabular}

    \clearpage
    \newpage

    \begin{tabular}{|l l l |}
    \hline
    Network & Network type & Original source \\
    \hline
    M\_PL\_037  & Pollination & \citep{Montero2005} \\
    M\_PL\_038  & Pollination & \citep{Montero2005} \\
    M\_PL\_039  & Pollination & \citep{Stald2003} \\
    M\_PL\_040  & Pollination & \citep{Ingversen2006} \\
    M\_PL\_041  & Pollination & \citep{Ingversen2006} \\
    M\_PL\_042  & Pollination & \citep{Philipp2006} \\
    M\_PL\_043  & Pollination & \citep{Montero2005} \\
    M\_PL\_044  & Pollination & \citep{Kato2000}  \\
    M\_PL\_045  & Pollination & \citep{Lundgren2005}  \\
    M\_PL\_046  & Pollination & \citep{Bundgaard2003} \\
    M\_PL\_047  & Pollination & \citep{Dupont2009a}  \\
    M\_PL\_048  & Pollination & \citep{Dupont2009a}  \\
    M\_PL\_049  & Pollination & \citep{Beks2006}  \\
    M\_PL\_050  & Pollination & \citep{Stald2003} \\
    M\_PL\_051  & Pollination & \citep{Vazquez2002} \\
    M\_PL\_052  & Pollination & \citep{Witt1998}  \\
    M\_PL\_053  & Pollination & \citep{Yamazaki2003}  \\
    M\_PL\_054  & Pollination & \citep{Kakutani1990}  \\
    M\_PL\_055  & Pollination & \citep{Kato1996}  \\
    M\_PL\_056  & Pollination & \citep{Kato1993}  \\
    M\_PL\_057  & Pollination & \citep{Inoue1990} \\
    M\_PL\_058  & Pollination & \citep{Bartomeus2008} \\
    M\_PL\_059  & Pollination & \citep{Bezerra2009} \\
    Basset  & Herbivory & \citep{Basset1999}  \\
    Bluthgen  & Herbivory & \citep{Bluthgen2006}  \\
    Bodner  & Herbivory & \citep{Bodner2010}  \\
    Coley & Herbivory & \citep{Coley2006} \\
    Ibanez  & Herbivory & \citep{Ibanez2013}  \\
    Joern\_altuda  & Herbivory & \citep{Joern1979} \\
    Joern\_marathon  & Herbivory & \citep{Joern1979} \\
    Novotny & Herbivory & \citep{Novotny2012} \\
    Peralta & Herbivory & \citep{Peralta2015} \\
    Sheldon & Herbivory & \citep{Sheldon1978} \\
    Ueckert & Herbivory & \citep{Ueckert1971} \\
    \hline
    \end{tabular}
    \end{table}

\clearpage
\newpage

\section*{Supporting information 2: Distributions of $p$-values for permuted networks}

% \begin{figure}[!h]
%     \begin{center}
%       \centerline{\includegraphics*[height=.62\textheight]{Figures/random_p_distributions.eps}}
%     \end{center}
%      \caption{Distributions of $p$-values obtained when comparing the strength of the relationship between niche overlap and phylogenetic distance in 999 permutations of each network in our dataset with 500 permutations of each permuted network. Each line in each panel represents the histogram of $p$-values for one network. Bins are 0.05 wide. Purple lines in panels A-L represent pollination networks while green lines in panels M-N represent plant-herbivore networks. Networks are as follows (darkest to lightest):
%      \textbf{A)} M_PL_001, M_PL_002, M_PL_003, M_PL_004, M_PL_005;
%      \textbf{B)} M_PL_006, M_PL_007, M_PL_008, M_PL_009, M_PL_010;
%      \textbf{C)} M_PL_011, M_PL_012, M_PL_013, M_PL_014, M_PL_015;
%      \textbf{D)} M_PL_016, M_PL_017, M_PL_018, M_PL_019, M_PL_020;
%      \textbf{E)} M_PL_021, M_PL_022, M_PL_023, M_PL_024, M_PL_025;
%      \textbf{F)} M_PL_026, M_PL_027, M_PL_028, M_PL_029, M_PL_030;
%      \textbf{G)} M_PL_031, M_PL_032, M_PL_033, M_PL_034, M_PL_035;
%      \textbf{H)} M_PL_036, M_PL_037, M_PL_038, M_PL_039, M_PL_040;     
%      \textbf{I)} M_PL_041, M_PL_042, M_PL_043, M_PL_044, M_PL_045;
%      \textbf{J)} M_PL_046, M_PL_047, M_PL_048, M_PL_049, M_PL_050;     
%      \textbf{K)} M_PL_051, M_PL_052, M_PL_053, M_PL_054, M_PL_055;
%      \textbf{L)} M_PL_056, M_PL_057, M_PL_058, M_PL_059;
%      \textbf{M)} Ibanez, Joern_altuda, Joern_marathon, Peralta, Sheldon;
%      \textbf{N)} Ueckert, Basset, Bluthgen, Bodner, Coley, Novotny.
%      For original sources for each network, see Table S1.
%      }
%     \label{within_family_regression}
%   \end{figure}

\end{spacing}

\clearpage

\newpage
\bibliographystyle{newphy}
\renewcommand*{\bibfont}{\raggedright}
\bibliography{noisn}

\end{document}