### 2019-05-23 manuscript/supporting_info.tex
checked bib, ready to send out revamped discussion

\documentclass[12pt]{article}  
\usepackage{amsmath}
\usepackage{url}
\usepackage[dvips]{graphicx}
\usepackage{multirow}
\usepackage{geometry}
\usepackage{pdflscape}
\usepackage{gensymb}
% \usepackage{rotating}
% make Figure 1 etc bold
\usepackage[labelfont=bf]{caption}
\usepackage{setspace}

\usepackage[running]{lineno}

\usepackage{dcolumn}
\newcolumntype{d}[1]{D{.}{.}{#1}}

%\usepackage{overcite}
\usepackage[round]{natbib}

\newcommand{\expect}[1]{\left\langle #1 \right\rangle}
\newcommand{\etal}{\textit{et al.\ }}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

% the abstract formatting
\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}
\renewcommand\refname{References}

% margin sizes`
\topmargin 0.0cm
\oddsidemargin 0.2cm

\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\title{Conservation of interaction partners between related plants varies widely across communities and between plant families - Supporting Information}

\author{Alyssa R. Cirtwill$^{1,2}$, Giulio V. Dalla Riva$^{3}$, Nick J. Baker$^{1}$,\\
Mikael Ohlsson$^{4}$, Isabelle Norstr\"{o}m$^{4}$, Inger-Marie Wohlfarth Hasle$^{4}$,\\
Joshua A. Thia$^{1,5}$, Christie J. Webber$^{1}$, Daniel B. Stouffer$^{1}$}
\date{\small$^1$Centre for Integrative Ecology, School of Biological Sciences\\
\medskip$^2$Department of Ecology,\\
Environment, and Plant Sciences (DEEP)\\
Stockholm University\\
114 19 Stockholm, Sweden\\
\medskip$^3$Biomathematics Research Centre, School of Mathematics and Statistics\\
University of Canterbury\\Private Bag 4800\\
Christchurch 8140, New Zealand\\
\medskip$^4$Department of Physics, Chemistry, and Biology (IFM)\\ Link\"{o}ping University\\ 581 83 Link\"{o}ping, Sweden\\
\medskip$^5$Present Address: School of Biological Sciences\\
University of Queensland\\Brisbane, QLD 4072, Australia }

\begin{document}
\maketitle
\baselineskip=8.5mm
\begin{spacing}{1.0}

\beginsupplement
\clearpage

\section*{Supporting information 1: Sources for networks}
  \begin{table}[!h]
    \caption{Original sources for each network used in our analyses.}
    \label{sources}
    \begin{center}
    \begin{tabular}{|l l m{10cm} |}
    \hline
    Network & Network type & Original source \\
    \hline
    M\_PL\_001  & Pollination & \citep{Arroyo1982}  \\
    M\_PL\_002  & Pollination & \citep{Arroyo1982}  \\
    M\_PL\_003  & Pollination & \citep{Arroyo1982}  \\
    M\_PL\_004  & Pollination & \citep{Barrett1987} \\
    M\_PL\_005  & Pollination & \citep{Clements1923}  \\
    M\_PL\_006  & Pollination & \citep{Dicks2002} \\
    M\_PL\_007  & Pollination & \citep{Dicks2002} \\
    M\_PL\_008  & Pollination & \citep{Dupont2003}  \\
    M\_PL\_009  & Pollination & \citep{Elberling1999} \\
    M\_PL\_010  & Pollination & Elberling, H. \& Olesen, J. M. Unpublished. \\
    M\_PL\_011  & Pollination & \citep{Olesen2002a}  \\
    M\_PL\_012  & Pollination & Olesen, J. M. Unpublished.  \\
    M\_PL\_013  & Pollination & \citep{Ollerton2003}  \\
    M\_PL\_014  & Pollination & \citep{Hocking1968} \\
    M\_PL\_015  & Pollination & \citep{Petanidou1991} \\
    M\_PL\_016  & Pollination & \citep{Herrera1988} \\
    M\_PL\_017  & Pollination & \citep{Memmott2002} \\
    M\_PL\_018  & Pollination & Olesen, J. M. Unpublished.  \\
    M\_PL\_019  & Pollination & \citep{Inouye1988}  \\
    M\_PL\_020  & Pollination & \citep{Kevan1970} \\
    M\_PL\_021  & Pollination & \citep{Kato1990}  \\
    M\_PL\_022  & Pollination & \citep{Medan2002} \\
    M\_PL\_023  & Pollination & \citep{Medan2002} \\
    M\_PL\_024  & Pollination & \citep{Mosquin1967} \\
    M\_PL\_025  & Pollination & \citep{Motten1982}  \\
    M\_PL\_026  & Pollination & \citep{McMullen1993}  \\
    M\_PL\_027  & Pollination & \citep{Primack1983} \\
    M\_PL\_028  & Pollination & \citep{Primack1983} \\
    M\_PL\_029  & Pollination & \citep{Primack1983} \\
    M\_PL\_030  & Pollination & \citep{Ramirez1992} \\
    M\_PL\_031  & Pollination & \citep{Ramirez1989} \\
    \hline
    \end{tabular}
    \end{center}
    \end{table}

    \clearpage
    \newpage

    \begin{table*}[h!]
    \begin{center}
    \begin{tabular}{|l l m{6cm} |}
    \hline
    Network & Network type & Original source \\
    \hline
    M\_PL\_032  & Pollination & \citep{Schemske1978}  \\
    M\_PL\_033  & Pollination & \citep{Small1976} \\
    M\_PL\_034  & Pollination & \citep{SmithRamirez2005}  \\
    M\_PL\_035  & Pollination & \citep{Percival1974}  \\
    M\_PL\_036  & Pollination & Olesen, J. M. Unpublished.  \\
    M\_PL\_037  & Pollination & \citep{Montero2005} \\
    M\_PL\_038  & Pollination & \citep{Montero2005} \\
    M\_PL\_039  & Pollination & \citep{Stald2003} \\
    M\_PL\_040  & Pollination & \citep{Ingversen2006} \\
    M\_PL\_041  & Pollination & \citep{Ingversen2006} \\
    M\_PL\_042  & Pollination & \citep{Philipp2006} \\
    M\_PL\_043  & Pollination & \citep{Montero2005} \\
    M\_PL\_044  & Pollination & \citep{Kato2000}  \\
    M\_PL\_045  & Pollination & \citep{Lundgren2005}  \\
    M\_PL\_046  & Pollination & \citep{Bundgaard2003} \\
    M\_PL\_047  & Pollination & \citep{Dupont2009a}  \\
    M\_PL\_048  & Pollination & \citep{Dupont2009a}  \\
    M\_PL\_049  & Pollination & \citep{Bek2006}  \\
    M\_PL\_050  & Pollination & \citep{Stald2003} \\
    M\_PL\_051  & Pollination & \citep{Vazquez2002} \\
    M\_PL\_052  & Pollination & \citep{Witt1998}  \\
    M\_PL\_053  & Pollination & \citep{Yamazaki2003}  \\
    M\_PL\_054  & Pollination & \citep{Kakutani1990}  \\
    M\_PL\_055  & Pollination & \citep{Kato1996}  \\
    M\_PL\_056  & Pollination & \citep{Kato1993}  \\
    M\_PL\_057  & Pollination & \citep{Inoue1990} \\
    M\_PL\_058  & Pollination & \citep{Bartomeus2008} \\
    M\_PL\_059  & Pollination & \citep{Bezerra2009} \\
    Basset  & Herbivory & \citep{Basset1996}  \\
    Bluthgen  & Herbivory & \citep{Bluthgen2006}  \\
    Bodner  & Herbivory & \citep{Bodner2010}  \\
    Coley & Herbivory & \citep{Coley2006} \\
    Ibanez  & Herbivory & \citep{Ibanez2013}  \\
    Joern\_altuda  & Herbivory & \citep{Joern1979} \\
    Joern\_marathon  & Herbivory & \citep{Joern1979} \\
    Novotny & Herbivory & \citep{Novotny2012} \\
    Peralta & Herbivory & \citep{Peralta2016} \\
    Sheldon & Herbivory & \citep{Sheldon1978} \\
    Ueckert & Herbivory & \citep{Ueckert1971} \\
    \hline
    \end{tabular}
    \end{center}
    \end{table*}
\clearpage
\newpage

\section*{Supporting information 2: Distributions of $p$-values for permuted networks}

    Comparing the permuted networks to permutations of the permuted networks, the slope obtained from the initial permuted network showed no clear relationship to the slopes obtained from 500 permutations of the permuted network. Averaged over the 1000 permutations of each observed network, the slope of the permuted network was more extreme than 48.1-51.3\% of the permutations of the permuted network. This confirms that shuffling phylogenetic distances between plant pairs destroys the relationship between distance and interaction partner overlap, and that further shuffling distances does not have a predictable effect.

\begin{figure}[!h]
    \begin{center}
      \centerline{\includegraphics*[width=.75\textwidth]{Figures/random_p_distributions.eps}}
    \end{center}
     \caption{\small Distributions of $p$-values obtained when comparing the strength of the relationship between niche overlap and phylogenetic distance in 999 permutations of each network in our dataset with 500 permutations of each permuted network. Each line in each panel represents the histogram of $p$-values for one network. Bins are 0.05 wide. Purple lines in panels A-L represent pollination networks while green lines in panels M-N represent plant-herbivore networks. A list of networks shown on each panel follows; see Table S1 for original sources.
     % Networks are as follows (darkest to lightest): \textbf{A)} M\_PL\_001, M\_PL\_002, M\_PL\_003, M\_PL\_004, M\_PL\_005;      \textbf{B)} M\_PL\_006, M\_PL\_007, M\_PL\_008, M\_PL\_009, M\_PL\_010;      \textbf{C)} M\_PL\_011, M\_PL\_012, M\_PL\_013, M\_PL\_014, M\_PL\_015;      \textbf{D)} M\_PL\_016, M\_PL\_017, M\_PL\_018, M\_PL\_019, M\_PL\_020;      \textbf{E)} M\_PL\_021, M\_PL\_022, M\_PL\_023, M\_PL\_024, M\_PL\_025;   \textbf{F)} M\_PL\_026, M\_PL\_027, M\_PL\_028, M\_PL\_029, M\_PL\_030;     \textbf{G)} M\_PL\_031, M\_PL\_032, M\_PL\_033, M\_PL\_034, M\_PL\_035;   \textbf{H)} M\_PL\_036, M\_PL\_037, M\_PL\_038, M\_PL\_039, M\_PL\_040;     \textbf{I)} M\_PL\_041, M\_PL\_042, M\_PL\_043, M\_PL\_044, M\_PL\_045; \textbf{J)} M\_PL\_046, M\_PL\_047, M\_PL\_048, M\_PL\_049, M\_PL\_050;      \textbf{K)} M\_PL\_051, M\_PL\_052, M\_PL\_053, M\_PL\_054, M\_PL\_055;     \textbf{L)} M\_PL\_056, M\_PL\_057, M\_PL\_058, M\_PL\_059;      \textbf{M)} Ibanez, Joern\_altuda, Joern\_marathon, Peralta, Sheldon;      \textbf{N)} Ueckert, Basset, Bluthgen, Bodner, Coley, Novotny.     For original sources for each network, see Table S1.
     }
    \label{within_family_regression}
  \end{figure}

\subsection*{Networks shown in each panel of Figure S1}

All names are as in Table S1. Networks included in each panel (from darkest to lightest line colours) are:

\noindent \textbf{A)} M\_PL\_001, M\_PL\_002, M\_PL\_003, M\_PL\_004, M\_PL\_005;      \\
\textbf{B)} M\_PL\_006, M\_PL\_007, M\_PL\_008, M\_PL\_009, M\_PL\_010;    \\
\textbf{C)} M\_PL\_011, M\_PL\_012, M\_PL\_013, M\_PL\_014, M\_PL\_015;  \\
\textbf{D)} M\_PL\_016, M\_PL\_017, M\_PL\_018, M\_PL\_019, M\_PL\_020;   \\
\textbf{E)} M\_PL\_021, M\_PL\_022, M\_PL\_023, M\_PL\_024, M\_PL\_025;  \\
\textbf{F)} M\_PL\_026, M\_PL\_027, M\_PL\_028, M\_PL\_029, M\_PL\_030;  \\
\textbf{G)} M\_PL\_031, M\_PL\_032, M\_PL\_033, M\_PL\_034, M\_PL\_035; \\
\textbf{H)} M\_PL\_036, M\_PL\_037, M\_PL\_038, M\_PL\_039, M\_PL\_040;   \\
\textbf{I)} M\_PL\_041, M\_PL\_042, M\_PL\_043, M\_PL\_044, M\_PL\_045;\\
\textbf{J)} M\_PL\_046, M\_PL\_047, M\_PL\_048, M\_PL\_049, M\_PL\_050;  \\
\textbf{K)} M\_PL\_051, M\_PL\_052, M\_PL\_053, M\_PL\_054, M\_PL\_055;   \\
\textbf{L)} M\_PL\_056, M\_PL\_057, M\_PL\_058, M\_PL\_059;  \\
\textbf{M)} Ibanez, Joern\_altuda, Joern\_marathon, Peralta, Sheldon;  \\
\textbf{N)} Ueckert, Basset, Bluthgen, Bodner, Coley, Novotny.

\end{spacing}

\clearpage

\newpage
\bibliographystyle{newphy}
\renewcommand*{\bibfont}{\raggedright}
\bibliography{manual}

\end{document}