### 2019-11-13 manuscript/reply_to_referees_2.tex
launching second reply

\documentclass[12pt]{letter}

%\usepackage[britdate]{stockholm-letter}
 \usepackage[britdate,alyssa-signature]{stockholm-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

%\usepackage{natbib}
% \bibpunct[; ]{(}{)}{;}{a}{,}{;}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{Postdoctoral fellow}
\department{Department of Ecology, Evolution, and Plant Sciences}
\location{Private Bag 4800}
\telephone{+64 3 364 2729}
\fax{+64 3 364 2590}
\email{alyssa.cirtwill@gmail.com}
\url{http://cirtwill.github.io}
\name{Alyssa R. Cirtwill}

\newcommand{\mytitle}{\emph{At a global scale, conservation of pollinators and herbivores between related plants varies widely across communities and between plant families.}}
\newcommand{\myjournal}{\emph{New Phytologist}}

\begin{document}

\begin{letter}{\bf Dr. Jana Vamosi\\
               Editor\\
               Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

\opening{Dear Dr. Vamosi:}

\closing{Regards,}

\end{letter}

\newpage

\setcounter{page}{1}

NPH-MS-2019-30803 At a global scale, conservation of pollinators and herbivores between related plants varies widely across communities and between plant families.
by Cirtwill, Alyssa; Dalla Riva, Giulio; Baker, Nick; Ohlsson, Mikael; Norström, Isabelle; Wohlfarth, Inger-Marie; Thia, Josh; Stouffer, Daniel

Dear Dr Cirtwill

I write you in regard to your submitted manuscript, NPH-MS-2019-30803 "At a global scale, conservation of pollinators and herbivores between related plants varies widely across communities and between plant families.”, which has now been reviewed. We found four reviewers (some original reviewers and some new) and they all generally enjoyed your revised manuscript. However, they also provided suggestions for some expanding certain sections and providing additional clarification. In particular, Reviewer #2 and #4 provide some detailed suggestions, especially with regard to keeping the aims and logic consistent, that will help strengthen the manuscript.

Although resolving these issues will require some restructuring, I think you should be able to handle these suggestions through revision (see below) and, ultimately, this manuscript will make a nice contribution to the literature.

Please carefully revise your manuscript and resubmit your revised version within 4 weeks. Please also include a summary of how you have incorporated the suggestions of the reviewers.

Thank you for submitting your work to New Phytologist.

Sincerely,

Jana Vamosi
Editor, New Phytologist

*****************************************************************
To revise your manuscript please click on the link below
*** PLEASE NOTE: This is a two-step process. After clicking on the link, you will be directed to a webpage to confirm. ***

https://mc.manuscriptcentral.com/newphytologist?URL_MASK=ebf0a265cb2d423bb35580e635386130

Alternatively log on to New Phytologist ScholarOne Manuscripts:
http://mc.manuscriptcentral.com/newphytologist

Enter your Author Centre.

Under the ‘My Manuscripts' list click on 'Manuscripts with Decisions'. From the list of manuscripts click the ‘create a revision' link, next to the manuscript you wish to submit a revision for. From here follow the instructions inputting you responses to the decision letter through to the uploading of your new files.

Please note you will be unable to make your revisions directly on the submitted ScholarOne Manuscripts version of your manuscript. Instead, you should revise your original word processor file, and then submit the revised version. When revising your manuscript, please prepare two versions: 1) the final revised manuscript and 2) the revised manuscript with revision marks showing. The latter file should be designated as a 'document with revision changes highlighted' when uploading.

In any correspondence regarding this manuscript, please include the manuscript reference number and copy the correspondence to New Phytologist Central Office (np-managinged@lancaster.ac.uk).
*************************************************************

Decision: Accept subject to revision

\clearpage

{\Large \bf Reply to Referee 1}

\begin{refquote}
Having read the manuscript and the response to previous reviewers’ comments, I do not see any further changes that need to be made. Any concerns I would have had, have been addressed by the latest round of revisions. I did not even spot any typos!
\end{refquote}

\textbf{R:} We thank the referee for taking the time to read our manuscript and for the compliment to our proof-reading.

\clearpage

{\Large \bf Reply to Referee 2}

	\begin{refquote}
	General comments

	The manuscript proposes to evaluate how phylogeny modulates species interaction patterns in plant-pollinator and a subset of plant-herbivory networks. In this regard, it explores an extremely relevant question for ecologists. However, it still needs work on the logic of the introduction section and methodological decisions need to be revised in order to achieve more confident results. The Introduction lacks focus and presents unclear aims, which leads to an unfocused discussion of the results, being one of the main weaknesses of the manuscript. Also, there is a lack of logical consistency between the abstract (methodologically-oriented) and the introduction and discussion sections (both biologically-oriented). On the other hand, we do not understand why authors do not incorporate ideas of species evolution and coevolution in a network context if you are trying to understand how phylogenetic constraints may modulate species interaction patterns in networks. Considering such exciting ideas may help to understand some of your results and may advance our understanding of  the importance evolutionary processes in determining species interaction patterns in communities.
	Another main point that you should revise across your manuscript is that you confound phylogenetic patterns with taxonomic patterns, for instance when thinking about congeneric species as closely related species (which is not necessarily true in several cases).

	\smallskip

	In what follows we share with authors our specific comments expecting that they will help to re-think the manuscript and get a stronger contribution.

	\end{refquote}

	The referee seems to view our basic question and analyses positively, but is rather critical of our  writing both in terms of logical clarity and style. We thank the referee for taking the time to provide such extensive feedback and have done our best to incorporate as many of their suggestions as possible. While we now mention the potential for coevolution to affect network structure, we have opted not to make this a major theme of the manuscript. What is and is not coevolution seems to be hotly debated. In the absence of any historical, paleontological, or genetic data to settle the question in our dataset, we feel it is not appropriate to make coevolutionary claims about our results. We address each comment in detail below (preceded by \textbf{R:}). It appears that the Referee wrote their specific comments before reading the manuscript in its entirety. Thus, some of the Referee's objections are in fact addressed at later points. In these cases, we have juxtaposed the Referee's comments to demonstrate this.

	1. General criticism of the language of the text

		\begin{refquote}

		Finally, the text needs English revision by the authors as we feel that some ideas are not clearly stated because of the way they are presented to readers.

		\end{refquote}

		\textbf{R:} We would like to point out that both the lead and corresponding authors of this manuscript are native English speakers and that Referees 1 and 3 praised the quality of our writing. Of course technical correctness does not necessarily imply clarity of expression, and so we have revised the manuscript according to the Referee's suggestions. [[Given the Referee's own moderate to poor command of English, this is a hell of a thing to say!!]]

	2. Title lacks clarity

		\begin{refquote}
		Specific comments
		Title
		Your title still lacks clarity and does not reflect the main message you state at the beginning of your discussion (in the first sentence of it you state that you found broad support for the hypothesis that more closely related plants tend to have more niche overlap). We feel that this reflects lack of logical consistency across your manuscript, so re-thinking it should help to define your title.
		\end{refquote}

	3. Summary too methodologically-oriented in general

		\begin{refquote}
			Summary
			General comments
			Your summary is too methodologically-oriented. In fact, when we were invited to be reviewers of your manuscript and received your abstract we thought that your paper represented a methodological contribution and it is not. After reading the comments that former reviewers sent you, your abstract seems to correspond to an older version of your manuscript. The thing is that by presenting an abstract that is methodologically-oriented when in fact you are not innovating in this sense, and having an introduction and a discussion that lack such methodological issue and are clearly more biologically-oriented can be read as that your work lacks focus.
		\end{refquote}

		\textbf{R:}

	4. Persistent mis-understanding of purpose and scope of within-family analyses

		\begin{refquote}
		Your abstract also evidences your confusion with phylogenetic issues and taxonomic ones. For instance, your hypothesis is that more closely related plants will share more pollinators and herbivores than less closely related. Then, why searching for different relationships within families? This decision is completely arbitrary is one of the weakest point of the manuscript in general.
		\end{refquote}

		\textbf{R:} Analyses of relationships between phylogenetic relatedness and niche overlap within families was specifically requested in a previous round of review [[check this]]. It also represents a major difference between our manuscript and previous work. We therefore disagree with the Referee that including this analysis is either "completely arbitrary" or "one of the weakest point(sic) of the manuscript in general". If more closely-related plants are indeed more likely to share interaction partners, then why shouldn't this be true within a family as well as between families? This additional level of detail demonstrates that not all plant families show the same relationships. We feel that this is useful. Based on point 6 (below), it appears that the Referee's objection is actually to the choice of family as our unit for detailed analysis rather than the analysis itself. We address this below.

		\begin{refquote}
		3rd paragraph:  which is your biological argument to consider families as the taxonomic level to search for differences across communities?
		\end{refquote}

		\textbf{R:} This comment suggests that the Referee was confused about our analyses. We did not search for differences between communities in our family-level analyses. Our between-community comparisons included comparisons between plant pairs pairs at all taxonomic levels. Our within-family analyses (based on pairs of species from the same community) were intended to detect differences between plant families \emph{across communities}. [[Clearly need to make this clearer in the summary, but R2 is also evidently a bit of an idiot. And probably Francophone. Maybe Indian?]]

		\begin{refquote}
		Aims
		Why considering the phylogenetic signal at the plant families level? There is no introduction for this aim. Why focusing at such taxonomic level?
		\end{refquote}

		\begin{refquote}
			Line 318. Within-family conservation of niche overlap. Your result is almost expected so we think is important to justify biologically why you decided to look at the family level.
		\end{refquote}

	5. Second paragraph of summary is too technical

		\begin{refquote}
		Your 2nd paragraph is too technical: which is the biological meaning of it? Why are you going to compare pollination and herbivory networks? Why your datasets are so unbalanced? Why are you considering a wide range of pollination networks and a narrow one for herbivory networks? Why you consider shared and unshared partners? Your first point in the abstract doesn't introduce a methodological problem to be solve. (...) "and whether this relationship varies with the composition of the plant community"? What does it mean?
		\end{refquote}

		\textbf{R:}

	6. Very confusing comment complaining about paragraph 4

		\begin{refquote}
		4st paragraph "Considering factors affecting the dominant plant families within
		a community may be the key to understanding the distribution of interactions." So, relationships are more functional? Why focusing on plant families? we are not sure the logic of your thoughts can be followed (can the reader reach this conclusion? is it ok? we mean, if there are other eco-evolutionary processes determining species interactions why focusing in the dominant plant families would be informative? I feel the way this last paragraph is written isn't clear enough and is your main paragraph.
		\end{refquote}

		\textbf{R:} This comment is difficult to follow. Part of the issue seems to be the Referee's mistaken belief that all of our analyses are at the level of plant families (addressed above). The remaining objection seems to be our suggestion that focusing on the plant groups which contribute the bulk of species in a community is more informative than focusing on species which represent unusual groups for the community. These dominant, species-rich groups contribute the bulk of unique interactions to a community, even if not necessarily the bulk of individuals. For this reason alone, it is worth understanding relationships between relatedness and niche overlaps in these groups. More pragmatically, species which are the lone representatives of a clade (we use families) do not provide much evidence for \emph{any} process.

	7. Complaint about keywords absent from the abstract

		\begin{refquote}
		Keywords
		Syndrome:  Why you do not mention such theory in your abstract?
		\end{refquote}

		\textbf{R:} The abstract/summary is under strict word limits, and we did not feel that inserting each keyword without a clear purpose could justify the elimination of other material. Given the Referee's confusion regarding our analyses, we are more convinced than ever that our decision to omit some keywords (which do appear in the introduction) in favour of spending more words explaining our methodology was correct.

	8. Mistake about the focus of the manuscript

		\begin{refquote}
		Introduction
		General. You focus on the compromise between pollination and herbivory interactions but then you do not use this approach in your analyses.
		\end{refquote}

		\textbf{R:} Potential tradeoffs between encouraging interactions with pollinators and discouraging herbivores are just one of the possible reasons why networks may not display the expected relationship between relatedness and niche overlap. Note that we also propose several others [[pull quotes]]. [[Apparently I need to adjust so it's clear that this isn't our main focus]]

	9. Suggestion to reframe the first paragraph in terms of functionality

		\begin{refquote}
		1st paragraph: We suggest this paragraph needs to be more functionally-oriented than phylogenetically-oriented. Syndromes may respond to the convergence of differently related organisms to a similar phenotype to attract the same pollinators. That is why plant syndromes can be more related with higher levels of taxonomic classification (orders) or to functionally defined groups like short and long-tongued bees, and not necessarily with sharing exactly the same pollinator species. Are you aware of articles showing that plant syndromes are phylogenetically conserved? We can think about many genera in which closely related plants show the same syndrome but bee or hummingbird syndromes are found within the family, so why should we expect that phylogeny has a strong signal? Or that more closely related plants will show more similar pollinators?
		\end{refquote}

		\textbf{R:}

	10. Referee hates paragraph 2 as well.

		\begin{refquote}
		2nd paragraph: your reasoning is not clear for us. Syndromes imply the convergence of phenotypic traits into one common set of traits because of their functionality in increasing fitness (because of the selection pressures imposed by functional groups of pollinators that can or cannot be closely related). Even if traits are heritable, if many unrelated groups converge into one functional group of interaction partners, why you may expect more similarity among closely related species? On the other hand, can we expect modular structures based on your reasoning? Then, how you explain the high incidence of nested structures in both mutualistic and antagonistic networks? In fact, what has been proposed for coevolution in mutualistic networks is the convergence of species traits to those of the species forming the core (see Guimaraes et al 2017, Nuismer et al. 2018). In antagonistic networks your line of reasoning may apply, but this will clearly depend on the intimacy of your network as most herbivory networks tend to show nested structures too. Your reasoning is based on pairwise coevolution and does not incorporate ideas of evolution and coevolution in networks, which may completely alter your predictions about niche overlap.
		\end{refquote}

		\begin{refquote}
		Lines 65-65. You wrote: Unrelated plants might also converge upon similar phenotypes, attracting a particularly efficient or abundant pollinator (Ollerton, 1996; Wilson et al., 2007; Ollerton et al., 2009; Ibanez et al., 2016). This is exactly our point: pollination syndromes may reflect trait convergence among unrelated species.
		\end{refquote}

		\textbf{R:} [[Fuck, this dude really seems to hate syndromes?]]

	11. Request to include interaction intimacy

		\begin{refquote}
		Lines 35-39. Maybe you should use the idea of interaction intimacy, since it clearly will determine the level of selective pressure landscape perceived by a given species?
		\end{refquote}

		\textbf{R:} We have no information on interaction strengths or intimacies, and so cannot test whether this is affecting our results. [[We could introduce the idea?]]

	12. Criticising methodological choices required by previous reviewers

		\begin{refquote}
		Methods
		Dataset
		Several networks can be reviewed in the literature or are available in different websites. Why authors decide not to consider several plant-herbivore networks but have no filter to separate plant-pollinator networks? The reason of why focusing in plant-pollinator and plant-herbivory networks is not clear for us as well why authors choose to select just one kind of herbivory interaction and as a consequence have a highly unbalanced design. Such decision seems to be crucial as it may modulate the results and conclusions they achieve. By eliminating the herbivory interactions that are more phylogenetically constrained their result that pollination interactions may be more conserved than herbivory interactions seems to be a by-product of such decision.
		\end{refquote}

		\textbf{R:} We restricted our selection of herbivory networks in response to a previous round of review. Without more details why (or if) the Referee believes that the herbivory networks we removed should be included, we have opted not to reject that earlier feedback. We would have appreciated specific tips for additional herbivory networks, but since the Referee does not provide any specifics there would appear to be little we can do to addrss this comment.

	13. Request to change "insects consuming leaves" to "chewing insects"

		\begin{refquote}
			Lines 93 to 96. We think that sap-sucking, leaf mining, and galling insects are also insect consuming leaves. Maybe you could change "insect consuming leaves" by "chewing insects"
		\end{refquote}

		\textbf{R:} [[Sure, why not.]]

	14. Mis-understanding of methods

		\begin{refquote}
			Measuring niche overlap
			Why you used absolute numbers to measure shared and unshared pollinators and not the Jaccard index? Or, instead, why you did not use a binomial regression considering the absolute number of the whole set of pollinators that can be shared ? You should use generalized linear models with a binomial distribution considering the total number of pollinators and shared pollinators to give more weight to those species sharing more pollinators in a larger sample (generalists s for instance). It is not clear for us if your analysis considers the full number of species over which you obtain the percentage of shared species.
		\end{refquote}

		\textbf{R:} If fact we did use a Jaccard index \emph{and} generalised linear models. Please read the methods carefully. These details can be found [[quote]]

	15. Confusing request to change methods

		\begin{refquote}
			176-178. Why you do not consider using plant species as another fixed factor to break the lack of independence? This is because generalist species may have more weight in your analyses than other species. It would be interesting to see how your results differ from an analysis considering such approach.
		\end{refquote}

		\textbf{R:} It sounds as though the Referee would like us to include plant species as a predictor in our regressions. Due to the fact that species generally appear in few networks (many in only a single network), this would result in near singularity. We would therefore be completely unable to observe any other patterns. If the Referee is referring to a potential effect of network size, we did in fact test this (lines XX-XX) [[[pull a quote]]]. If the Referee did not mean either of these possibilities, we apologise but are unable to understand this comment.

	16. Un-justified request to include connectance

		\begin{refquote}
			198-209. Why did you decide to control just for network size and do not consider network connectance? You provide methodological information to explain your decision but there is no biological argument explaining why network size may influence the relationship between phylogenetic relatedness and niche overlap.
		\end{refquote}

		\textbf{R:} We did not include connectance because we can offer no justification (biological or methodological) why the overall connectance of the network should affect the strength of the relationship between relatedness and niche overlap. It is possible that there might be an effect of connectance on the intercept of this relationship (as we would expect higher overlap in general when connectance is high), but we see no particular reason why connectance should change how closely-related species differ from distantly-related species in the same network. Since the Referee also does not provide any reasoning behind this request, we have opted to leave our analyses as-is.

		[[Maybe I can add a bit of a biological argument that network size can affect overlap? I'm not sure what methodological but not biological means here.]]

	17. Poor understanding of methods

		\begin{refquote}
			Line 210 Linking network level trends and community composition

			(related with this part of the methods but referring to the Introduction section) Which are the hypothesis for this analysis? How do you think that family composition (and no species composition) could affect the relationship between phylogenetic  distances and niche overlap?
			In the analysis it is not clear for us which are the dependent and independent variables. Is the independent variable a matrix with coefficients of similarity between networks?  Do you want to explore if more similar networks have a more similar relationship (slope) between phylogenetic distance and niche overlap? Or you want to investigate if the slopes are more o less strong depending on the local composition of communities (for each network separately) ? If this is the case you could compare slopes with a local diversity index (e.g. Shannon) instead of a beta diversity index.
		\end{refquote}

		\textbf{R:}

	18. Condescending critique of manuscript structure

		\begin{refquote}
			Results
			General: Please, do not include discussion in your results, it may confound readers.
		\end{refquote}

		\textbf{R:} The Referee appears to be confounded by our manuscript in general and is extremely dismissive of our understanding of how to write a manuscript...

	19. Possible complaint about mixing numbers and percentages?

		\begin{refquote}
			Line 262. Your first sentence of results do not correspond with what follows. We will suggest you use most networks, or most herbivory networks and half pollination networks, or using the percentages you present in the discussion section.
		\end{refquote}

		\textbf{R:} This comment is in such poor English as to be nearly unintelligible. It is not clear whether the Referee is suggesting that we repeat our analyses using half of the pollination networks (without any clear reason as to why) or complaining that we have mixed qualitative descriptors, numbers, and percentages. We assume the latter as we are able to address this complaint. [[check what's going on]]

	20. Confusion about methodology

		\begin{refquote}
			Line 310. How was community composition measured? Is difficult to evaluate the quality of this result because you haven't described this methodology.
		\end{refquote}

		\textbf{R:} Community composition is the set of plant species that make up a community. It was defined by the researchers constructing each network according to the specific methodology employed in each case. [[Add a parenthetical definition]]

	21. Confusion about permutation methodology

		\begin{refquote}
			Minor comments on Results
			297-298 You permuted interactions or phylogenetic distances?
		\end{refquote}

		\textbf{R:} [[double-check and provide a line ref]]

	22. Request to change direction of relationship?
		\begin{refquote}
			Fig 1. "...the probability of a pair of plants sharing an interaction partner increased with increasing phylogenetic distance..." please, write decreased instead of increased.
		\end{refquote}

		\textbf{R:} Sweet Satan, I hope this isn't a typo?

	23. Difference between minimum and maximum slopes not visible

		\begin{refquote}
			Fig 2. It is no clear for us. Minimum and maximum slopes obtained from 999 permutations are the same value?
		\end{refquote}

		\textbf{R:} [[Check that caption states that min and max are very close together since permutation completely destroyed pattern.]]

	24. Missing word in line 328

		\begin{refquote}
			Line  328: "...increasing phylogenetic distance in 12...", what is 12? the number of families?
		\end{refquote}

		\textbf{R:}

	25. Sundry complaints about table numbering and figure referencing

		\begin{refquote}
			Table number seem to be erroneous as you first mention table 2 and then table 1.

			\smallskip

			Figure 3 is only referenced in the text when talking about herbivory networks but not for pollination networks.
		\end{refquote}

		\textbf{R:}

	26. Blanket demand for more evolution content

		\begin{refquote}
			Discussion
			General
			We suggest you include ideas about evolution and coevolution in networks in your discussion. It is important to recognize that network structure is in part a consequence of the evolutionary process and that we cannot think in the pairwise-coevolutionary framework as the only one modulating trait evolution in species rich assemblages.
			You state that the fact that the proportion of specialist species may explain differences between plant-pollination and plant-herbivore network patterns and you discuss the mathematical consequences of such difference but you should discuss the ecological and evolutionary causes and consequences of having higher specialization in pollination networks. This is also related with your methodological decision of excluding more specialized herbivores, isn't it?
		\end{refquote}

		\textbf{R:} THIS IS NOT AN EVOLUTION PAPER! WE HAVE NO DATA TO INFER ANYTHING ABOUT THE EVOLUTION OF THESE SYSTEMS!

	27. Debate over the meaning of "broad"

		\begin{refquote}
			338-339. You state We found broad support for the hypothesis that more closely-related pairs of plants have a higher degree of niche overlap. This is a very strong sentence that is not supported by your results. You found evidence in 55% of plant-pollination networks (almost half of your dataset). In fact, it is interesting to note that even the slope is more pronounced in mutualistic networks you found higher support in herbivory networks, as a higher percentage of networks showed a tendency.
		\end{refquote}

		\textbf{R:}

	28. Another complaint about sub-setting the herbivory networks, possibly addressable in text

		\begin{refquote}
			351-352 The herbivory networks did not contain as many obligate specialists, but we note that many herbivorous insects are oligotrophs which consume only a few closely-related hosts (Novotny & Basset, 2005; Yguel et al., 2011). This sentence is confuse because you have a subsample of herbivory networks and then this seems to be a by-product of your study design associated with your decision of removing networks that may show strong phylogenetic signal.
		\end{refquote}

		\textbf{R:}

	29. Claims that congeneric species are not closely related

		\begin{refquote}
			357-359 This suggests that conservation of interaction partners among closely related plants (e.g., congeners or members of the same subfamilies) is more important than phylogenetic signal from deeper within the phylogenetic tree. Congeneric species can be not closely related, please revise sentences like this one across the manuscript because you are confounding phylogeny with taxonomy.
		\end{refquote}

		\textbf{R:} Does the Referee really mean to suggest that plants in the same genus are less closely-related than plants in different families? Without some sources to substantiate this, we find this claim very far-fetched!

	30. Another demand to discuss intimacy

		\begin{refquote}
			361-365. How interaction intimacy may explain your results? Please, discuss your results in light of the interplay between interaction type (pollination, herbivory) and intimacy (see Hembry et al. 2018).
		\end{refquote}

		\textbf{R:}

\clearpage

{\Large \bf Reply to Referee 3}

	\begin{refquote}
		The study is well designed and the manuscript well written, and the authors clearly demonstrate their comprehensive grasp of the subject and the applied statistical analyses.
		My only more general comment is whether authors have considered potential applications of such phylogenetic approaches for ecological restoration? If closely-related species provide similar functions in ecological networks, they could be used interchangeably to restore ecological interactions/functions. This is an exciting potential application of phylogenetic data in biodiversity conservation and restoration, and could be touched upon/discussed in this manuscript.. 
		Otherwise, I have only minor comments.
	\end{refquote}

	\textbf{R:} We thank the Referee for their compliments to our approach and to our writing, as well as their more detailed comments (addressed one-by one below; replies preceded by \textbf{R:}). We had not considered this potential application for ecological restoration but are very pleased to have it pointed out. It is often difficult to see how theoretical work may relate to conservation practice, so we are glad to take the opportunity to make a link here. [[Add a bit of this, I think we have room]]

	1. Title is confusing

		\begin{refquote}
			Title is confusing – difficult to pick out key message. And given that your analyses broadly give support for effects of phylogenetic distance on niche overlap, why not focus on this message, rather than variability at among individual networks/plant families
		\end{refquote}

		\textbf{R:}

	2. Request to state that networks are qualitative rather than quantitative

	3. L298 – Please be careful with use of “lesser/greater” “more negative or positive than expected from permuted networks”

	4. Confusion about result in line 322

		\begin{refquote}
			L322 – how was this result obtained if insect phylogenies were not caclculated? Do you mean “more closely-related plants in pollinator networks”?
		\end{refquote}

		\textbf{R:}

\clearpage

{\Large \bf Reply to Referee 4}

	\begin{refquote}
		Understanding what drives the structure of networks of interactions is key to predict the structure of plant-insect communities (e.g., for conservation or restoration purposes, or to support benefit organisms in agriculture).
		The present study focuses on the contribution of the evolutionary history to the interaction structure, and makes a significant contribution to the understanding of the variability of the phylogenetic signal in plant-insect networks. Here, the strength of the phylogenetic signal is actually quantified while it usually relies on correlation scores, and the effect of plant composition and the identity of plant families are examined to assess whether they could explain the variability of trends described in the literature.
		The dense writing turns out to be a double-edge sword: a lot of information on drivers of phylogenetic signals is compiled in here, but the text is sometime hard to follow. Hence, my following comments mostly pertains to the form, and I trust the authors should be able to address them fairly easily.
	\end{refquote}

	\textbf{R:} We thank the Referee for their compliments to our analysis and for taking the time to review our manuscript. We appreciate that our writing can be dense at times and appreciate the Referee pointing out places where we may clarify our prose. Below, we respond to each comment in detail (preceded by \textbf{R:}). We are confident that revising the manuscript in response to the Referee's comments has improved the manuscript.

This is the first time I review the present manuscript, so I apologise in advance if my comments were already addressed in previous reviewing rounds (despite checking with the most recent reply to reviewers).

Statistical analyses

* Modelling the relationship between niche overlap and phylogenetic distance
I did not managed to fully grasp the statistical analysis right away (I reckon I'm not a statistical master, but this may be the case of other readers), and I think it was mostly due to the way the methods are described. The writing is nice, but some tiny details are missing while they could help the reader to understand more quickly how the authors distinguish the use of the {M_ij, U_ij}-tuple from that of J_ij.
Based on l. 143-167, I understand that the number of shared interactions M_ij is modelled with a binomial distribution, with the number of trials being M_ij+U_ij, and the probability of success being w_ij (M_ij ~ B(M_ij + U_ij, w_ij)). The {M_ij, U_ij}-tuple would appear in R as follows:
model ← glm(cbind(M_ij, M_ij+U_ij) ~ d_ij + ..., family = “binomial”)
But, can we actually say that there are two dependent variables? To me, it is the number of shared interactions that is modelled as a binomial process, not both the number of shared and unshared interactions.

I suppose that the analysis described in paragraph starting from l. 168 corresponds to actually modelling the Jaccard index (J_ij ~ ...). If this is right, I think this should be specified in the second paragraph (no direct mention of the Jaccard index is currently made, except the proportion l. 170 between brackets although it is not exactly written as in eq. (1)), while in the first paragraph, an equation should specify that M_ij ~ B(M_ij + U_ij, w_ij)).

Another point pertains to the definition of “trials” in the binomial-distributed process. If I get it right, these correspond to the union of species i and j interactions (of size P_i + P_j -2M_ij = M_ij + U_ij). However, these are not independent for reasons highlighted in the introduction (e.g., competition, facilitation between plants and/or insects). This may question the robustness of the analysis, but testing the effect of the plant community composition (negative) allows to verify this. This should be specified, I think.

* Evaluating type I and type II errors: Permutations of permutations
Paragraph p. 10, l. 187-197 is a bit confusing because of phrasings such as “permutations/permuting of the permuted networks”. I don't have a better way to name this second round of permutations, but the overall is difficult to follow and could be shortened while still conveying the same message.
Actually, I find the comment in Supplementary information 3 clearer, although no reference to Supplementary information 3 is made.
Besides, as slopes of regressions in permuted networks are tightly grouped around zero (Fig. 2, thin lines indicating extrema of slopes distribution overlapping each other), so type II errors do not seem to be a big problem here.

* Testing the effect of network size on the existence of a phylogenetic signal
Here, it is specifically the number of plant pairs that is studied. I wonder whether the phylogenetic diversity of plants would not be more adapted here. Is it overlooked because similar among networks?
I think knowing more about this effect would further illuminate why the strength of the phylogenetic signal varies between networks and would be complementary to the analysis testing the effect of plant community composition.

* Testing within-family existence of a phylogenetic signal
Examining the strength of phylogenetic signal within plant families is an interesting idea. However, this leads to multiple testing which is prone to type I errors and I could not find mention of correction for this multiple testing in the text. There are actually high chances that there is at least one significant test in each table (1 – (1 – alpha)^N = 0.36 and 0.91 respectively, alpha = 0.05). Yet, the authors can indicate that there are low chances to have that many significant phylogenetic signals (using the probability mass function of the Binomial distribution, see Moran 2003 in Oikos).

On the effect of specialist pollinators

I think that Reviewer #3 made an interesting remark on specialist pollinators, and I agree with the authors on the impossibility to assess sampling completeness on the network of this database. However, I think this should be mentioned in the text, at least to invite future works to provide their sampling effort when publishing new empirical data (hence, helping to improve future analyses).

On plant families involved in different types of interactions

p. 19-20, l. 432-448: This paragraph on systems for which both pollination and herbivory were sampled is a bit fuzzy. Talking about how diverse types of interactions can work together is a nice way to bring back the study to a broader context, though.
First, I think the sentence of simultaneous selection pressure of pollination and herbivory could be a bit clarified (l. 439-442, “Plants may not be able to respond (…).”). Strauss & Whitthall’s chapter in the book “Ecology and evolution of flowers” gives a good insight on how these selection pressures can act together or antagonistically in shaping plants phenotypes (and eventually interactions, although not specifically discussed in their chapter).
Second, the structure of interaction patterns could be discussed with more references than Sauve et al. (2016). Astegiano et al. (2017) make an interesting point on the asymmetry of interaction patterns in such systems (using Pearse and Altermatt (2011)’s data on Lepidotera diet, and citing the Norwood farm data set of Pocock et al. 2012, and the Donana network in Melian et al. 2009).
Finally, the existence dynamical drivers of the structure of pollination and herbivory is mentioned. Sauve et al. (2014, 2016) indeed suggest that the way pollination and herbivory are distributed on the plant community may matter for community stability, and they studied one case where this structure is indeed not random. Yet, they do not say that these communities should be more stable (compared to which ones since plants undergo both herbivory and pollination in nature?). Whether the interconnection structure arises because it provides greater stability (as discussed by Sauve et al. 2016), or as a by-product of the assembly process (Maynard et al. 2018, Ecol. Lett.) remains to be discussed for networks combining different types of networks.
Despite the length of this last comment, I don’t mean that the authors should develop proportionally this paragraph, but rather sharpen it.

Minor comments

In the abstract, item 2 « shared and unshared interactions »: if it does not breach the word limit, I think a few words specifying why this is important would better showcase the study, and help to understand how it contributes to the study novelty.

p.3, l. 1-2, citation of Sauve et al. 2016 (Ecology) to say that interactions between plants and animals are critical to plants’ life cycle. I’m not sure this is the best pick as Sauve et al. rather look at how pollination and herbivory interactions are distributed in the plant community and how it may affect community stability. The following references are adequate though.

p. 3, l. 9-10 « A plant’s traits are also (...) », I think there is a typo here. Shouldn’t « a » be removed ?

p. 3, l. 22-25 : There is a lot of citations here. I agree they all point to « mixed results » but this pertains to different aspects of phylogenetic signal. Some highlight differences between mutualistic and antagonistic systems, other rather focus on difference of phylogenetic signal between trophic levels, or even focus on subset of interaction networks. I think, it would be more relevant to the reader to cite these different works to explain how mixed results are (as done in the text following this long citation list).

p. 5, l. 69-71 « In either case, dissimilarity (…). » This sentence helped me to understand why it is important to look at unshared interactions. To be better showcased ?

p. 5, l. 78 « a pairwise perspective » : This contributes to the manuscript novelty. Maybe this should be further specified?

p. 6, l. 89-90 « These networks span a range of biomes (…). » In supporting information, the countries are not indicated in the list. Alternatively, and to emphasize on the title following Reviewer #2 suggestion in the previous round of review, a figure showing the spatial distributions of the 59 + 11 network on a world map would be welcome here.

p. 7, l. 125-141 : I think the beginning of this section should be more straightforward and contrast better the information provided by M_ij and U_ij (starting with them) with that of the Jaccard index J_ij. In addition, J_ij is never mentioned again as such (but is an estimate of w_ij).

p. 14, l. 297-298 “Comparing the results in the observed networks with those obtained after permuting phylogenetic distances across pairs of plants (...)”: In the methods, interactions are shuffled, not phylogenetic distances. Doing one of the other seems equivalent for this test, but this rephrasing is a bit confusing.

p. 15, l. 322-324 “ More closely-related pollinators did, however, tend to share fewer interaction partners (…).”: This is the first time that conservatism of interactions is checked from the pollinators perspective in the manuscript. Is it a typo or did I miss something in the manuscript?

Legend of Fig. 3 “(…) Melastomatoaceae and Poaceae in pollination networks are not significantly different from zero.” but Table 2 shows the opposite for Poaceae.

In the figures, \beta_distance is alternatively named “slope of regression line” (Fig. 2) and “change in log odds” (Fig. 3).

References

Altermatt, F. and Pearse, I.S., 2011. Similarity and specialization of the larval versus adult diet of European butterflies and moths. The American Naturalist, 178(3), pp.372-382.

Astegiano, J., Altermatt, F. and Massol, F., 2017. Disentangling the co-structure of multilayer interaction networks: degree distribution and module composition in two-layer bipartite networks. Scientific reports, 7(1), p.15465.

Maynard, D.S., Serván, C.A. and Allesina, S., 2018. Network spandrels reflect ecological assembly. Ecology letters, 21(3), pp.324-334.

Melián, C.J., Bascompte, J., Jordano, P. and Krivan, V., 2009. Diversity in a complex ecological network with two interaction types. Oikos, 118(1), pp.122-130.

Moran, M.D., 2003. Arguments for rejecting the sequential Bonferroni in ecological studies. Oikos, 100(2), pp.403-405.

Strauss, S.Y. and Whittall, J.B., 2006. Non-pollinator agents of selection on floral traits. Ecology and evolution of flowers, pp.120-138.

  \newpage

\bibliographystyle{prsb} 
\bibliography{abbreviated}

\end{document}